Interactive Graphics for Visually Diagnosing Forest Classifiers in R

### 2022-06-30 response_reviewers.Rmd

Thank you very much.

Best regards,

Springer Journals Editorial Office
%JOURNALFULLTITLE

-->

\textcolor{violet}{Thank you. We appreciate the chance to do a further revision, and hope that these changes are satisfactory.}

@@

The paper and its accompanying app seem to be a good fit for Computational Statistics, as they apply interesting visualization techniques to a tree-based ensemble model that will surely prove useful for practitioners.

However, for this potential to materialize the authors should put more emphasis on the limitations of their approach. In particular, the app focuses on the first three nodes of the trees in question, which would be much less informative in settings requiring deep trees. \textcolor{violet}{The app was modified to adress this issue. In the new application the user can select specific nodes to see (from one to all the nodes). If there are a lot of nodes will be better to select 4 or less each time to see better the complete display but the application is not limited to see just the first three nodes.}

Furthermore, plots such as the parallel coordinate plot would lose their appeal in settings with thousands of variables. A brief discussion of circumstances limiting the usability of the app would address these concerns and allow users to easily determine whether the app is suitable for their application.

\textcolor{violet}{The discussion has been extended to include limitations.}

### 2022-06-30 PPforestVizupdate.Rnw
modifications to response
revised Discussion

\begin{document}

@@

The app is implemented with new technology for interactive graphics provided by the \verb# plotly# package. It is one of the first uses of these new tools.

One challenge to use plotly is that  when layers with different data are created in a ggplot2,  it is difficult to  specify the unique keys required for linking with another plot.

The application have some limitations that need to be pointed out. If the number of observation is big, the overploting will be an issue in some of the selected visualizations, and if the number of variables increases, plots like parallel coordinate plot and importance variable plot should include the option to select some variables each time or a different approach must be considered.

There are many possible extensions to the app that could help to become a tool for model refinement: (1) Using the diagnostics to weed out under-performing models in the ensemble; (2) Identifying and boosting models that perform well, particularly if they do well for problematic subsets of the data; (3) Problematic cases could be removed, and ensembles re-fit; (4) Classes as a whole could be aggregated or re-organised as suggested by the model diagnostics, to produce a more effective hierarchical approach to the multiple class problem. Working within the R environment makes all of these desires available using command line outside the app, given the unique IDs of models and cases can be exported from the app.

The app has helped to identify ways to improve the  PPtree algorithm and consequently, the PPF model which especially apply to multiclass problems. Multiple splits for the same class would enable nonlinear classifications. Split criteria tend to place boundaries too close to some groups, due to heteroskedasticity being induced by aggregating classes. Forests are not always better than their constituent trees, and if the trees can be built better, the forest will provide stronger predictions.

>>>>

\begin{document}
\SweaveOpts{concordance=TRUE}

@@

The app is implemented using the technology for interactive graphics provided by the \verb# plotly# package. It was one of the first to experiment with the application of these tools. One challenge with plotly is that  when layers containing different data are created in a ggplot2, it is difficult to specify the unique keys required for linking with other plots.

There are some limitations with providing functionality through the app, and generally with the displays provided. Successful apps make it easier for the user to do common tasks, and this is the design principal here. Only limited functionality is provided to the user, but the code is available and a knowledgeable data analyst might use this as a basis for extending the methods as appropriate for theit data. Performance of the app is good for relatively small data sets. If the number of observation is large, overploting will be an issue in some of the visualizations. If the number of variables is large, the parallel coordinate plot and plot of variable importance would benefit by the addition of a user input to select variables.

There are many possible extensions to the app that could help it to be a tool for model refinement:

\begin{enumerate}
\item Using the diagnostics to weed out under-performing models in the ensemble. 
\item Identifying and boosting models that perform well, particularly if they do well for problematic subsets of the data.
\item Problematic cases could be removed, and ensembles re-fit. 
\item Classes as a whole could be aggregated or re-organised as suggested by the model diagnostics, to produce a more effective hierarchical approach to the multiple class problem.

\noindent Working within the R environment makes all of these possible when using command line outside the app, given that the unique IDs of models and cases can be exported from the app.

The app has helped to identify ways to improve the PPtree algorithm and consequently, the PPF model which especially apply to multiclass problems. Multiple splits for the same class would enable nonlinear classifications. Split criteria tend to place boundaries too close to some groups, due to heteroskedasticity being induced by aggregating classes. Forests are not always better than their constituent trees, and if the trees can be built better, the forest will provide better predictions.
